Revisiting the hydrodynamic modulation of short waves by longer waves

### 2024-09-26 paper/paper.tex
Near final edits

\begin{equation}
\label{eq:lhs1960wavenumber}
\widetilde{k} = k (1 + \varepsilon_L)
\end{equation}
where $\widetilde{k}$ and $k$ are the modulated and unmodulated short-wave wavenumbers,
respectively, and $\varepsilon_L = a_L k_L$ is the steepness of the long wave with
wavenumber $k_L$ and amplitude $a_L$.
Hydrodynamic modulation is a key process for the development and
interpretation of remote sensing of ocean surface waves and currents
\citep{keller1975microwave,hara1994hydrodynamic}, and is distinguished
from aerodynamic modulation \citep{donelan1987effect,belcher1999wave,chen2000effects}.
\citet{phillips1981dispersion} and \citet{longuet1987propagation} revisited the
hydrodynamic modulation problem by considering nonlinear long waves and
variations in the effective gravity of short waves.
Their results yielded significantly stronger modulation than previously predicted.
\citet{henyey1988energy} derived an analytical solution for the modulation of
short waves by long waves using Hamiltonian mechanics and reported the modulation
of similar magnitude to that of \citet{longuet1987propagation}.
\citet{zhang1990evolution} considered weakly nonlinear short waves
using a non-linear Schr\" odinger equation and found stronger wavenumber
modulation but weaker amplitude modulation than those predicted by
\citet{longuet1987propagation}.
The variety of analytical and numerical frameworks (Eulerian, Hamiltonian,
Schr\" odinger) that relatively closely reproduce steady solutions for short-wave
modulation suggests that the solutions are robust.
Alternative analytical and numerical approaches to the modulation problem remain
of interest, especially if higher degrees of nonlinearity can be considered
using simpler approaches.

More recently, \citet{peureux2021unsteady} asked whether the steady solutions
of short-wave modulation are realistic by simulating the full wave conservation
equations.
They found that the short waves grow unsteadily due to the propagation of
longer waves, suggesting that the steady-state solutions may only be valid for
a few long-wave periods before the short waves begin to break due to excessive
steepness.
However, the unsteadiness of their solutions occurs in a specific scenario of
a uniform short-wave field that suddenly appears in a uniform long-wave train,
and the numerical simulations stabilize if the short-wave field is initialized
from existing steady solutions.
Nevertheless, their results do put into question whether the short-wave
modulation predicted by theory is generally steady over many long-wave periods
or not.

In this paper, we revisit this problem from linear wave theory and
derive alternative non-linear solutions for the modulation of short waves,
evaluated at the long-wave surface.
The analytical solutions yield modulations that are stronger than the original solution
by \citet{longuet1960changes}, but similar to the nonlinear numerical solutions
by \citet{longuet1987propagation} and \citet{zhang1990evolution} in long-wave
steepness up to $\approx 0.2$.
We also perform numerical simulations of the full wave crest and action
conservation laws and evaluate the effect of long-wave groups on the short waves.
The numerical simulations are used to determine the validity of the approximate
steady solution and to investigate the unsteadiness of the short-wave modulation
found by \citet{peureux2021unsteady}.
The main prior studies on hydrodynamic modulation, and their assumptions and
approaches to the solution, are summarized in Table
\ref{table:modulation_literature_summary}.

\begin{table}
\begin{center}
\def~{\hphantom{0}}
\begin{tabular}{cccccc}
Reference   & Governing & Gravity & Nonlinear  & Nonlinear   & Solution \\
            & equations & varies  & long waves & short waves &          \\
\hline
L-HS 1960   & Velocity potential  & No  & No  & No  & Analytical \\
\hline
P 1981      & Crest and action    & Yes & Yes & No  & Numerical \\
            & conservation        &     &     &     &           \\
\hline
L-H 1987    & Velocity potential  & Yes & Yes & No  & Numerical \\
\hline
HCDSW 1988  & Hamiltonian         & Yes & Yes & No  & Analytical \\
\hline
ZM 1990     & Schr\" odinger      & Yes & Yes & Yes & Numerical \\
\hline
PAG 2021    & Crest and action    & Yes & No  & No  & Numerical \\
            & conservation        &     &     &     &           \\
\hline
This paper  & Crest and action    & Yes & Yes & No  & Analytical \\
            & conservation        &     &     &     & and numerical \\
\hline
\end{tabular}
\caption{
  Summary of prior
  \citep{longuet1960changes,phillips1981dispersion,longuet1987propagation,henyey1988energy,zhang1990evolution,peureux2021unsteady}
  and present (this paper) approaches to calculating the modulation of short
  waves by long waves.
}
\label{table:modulation_literature_summary}
\end{center}
\end{table}

This paper first presents the governing equations and then the analytical
solutions for the modulation of short waves by long waves and compares them to
the original solution by \citet{longuet1960changes}.
Then, we compare the analytical solutions to numerical solutions of the full
wave conservation equations in section \ref{section:numerical_solutions}.
In section \ref{section:scale_analysis} we perform a scale analysis of the
analytical solutions to quantify the relative importance of the terms in the
full equation set.
Finally, we summarize the results and discuss their implications in section
\ref{section:conclusions}.

@@

We consider inviscid, irrotational, and incompressible flow.
In these conditions the linear (sine-like), deep-water, surface gravity waves
obey the dispersion relation:

\begin{equation}
\label{eq:dispersion}
\omega = \sqrt{gk} + k U
\end{equation}
where $\omega$ is the angular frequency, $g$ is the gravitational acceleration,
$k$ is the wavenumber, and $U$ is the mean advective current in the direction
of the wave propagation.
A current in the direction of the waves increases its absolute frequency,
while the wavenumber remains unchanged.
An important limitation here is that $U$ must be slowly varying on the scales of
the wave period \citep{bretherton1968wavetrains}.
In the context of short waves riding on long waves, the mean advective current
is the horizontal near-surface orbital velocity of the long wave.
The evolution of the short-wave wavenumber is described by the conservation of
wave crests \citep{phillips1981dispersion}:

\begin{equation}
\label{eq:wave_crests}
\dfrac{\partial k}{\partial t}
+ \dfrac{\partial \omega}{\partial x}
= 0
\end{equation}
where $t$ and $x$ are time and space, respectively.
This relation states that the wavenumber must change when the frequency varies
in space.

Equations (\ref{eq:dispersion}) and (\ref{eq:wave_crests}) only describe the
evolution of the short-wave wavenumber.
To describe the evolution of the wave amplitude, we use the conservation of wave
action in absence of sources (e.g. due to wind) and sinks (e.g. due to whitecapping):

\begin{equation}
\label{eq:wave_action}
\dfrac{\partial N}{\partial t}
+ \dfrac{\partial}{\partial x} \left[\left(C_g + U\right)N\right]
= 0
\end{equation}
where $N$ is the action density, and $C_g = \partial \omega / \partial k$ is the
group speed.
In deep water, the group speed is half the phase speed, so $C_g = 1/2\sqrt{g/k}$.
Although the absence of wave growth and dissipation is not generally realistic
in the open ocean, it is a helpful approximation that allows us to isolate the
hydrodynamic modulation effects solely due to the motion of the long waves.

Short waves that ride on the surface of the longer waves move in an accelerated
reference frame and thus experience effective gravitational acceleration that
changes in space and time \citep{phillips1981dispersion,longuet1986eulerian,longuet1987propagation}.
From Eqs. (\ref{eq:dispersion}) and (\ref{eq:wave_action}) we obtain:

\begin{equation}
\label{eq:wavenumber}
\dfrac{\partial k}{\partial t}
+ \left(C_g + U\right) \dfrac{\partial k}{\partial x}
+ k \dfrac{\partial U}{\partial x}
+ \dfrac{1}{2} \sqrt{\dfrac{k}{g}} \dfrac{\partial g}{\partial x}
= 0
\end{equation}
From left to right, the terms in this equation represent the change in wavenumber
due to the propagation and advection by ambient current $U$, the convergence of
the ambient current, and the spatial inhomogeneity of the gravitational acceleration.
In a non-accelerated reference frame (i.e. in absence of longer waves), the last
term vanishes.
Otherwise, this term is necessary to conserve the wave crests.

Similar to Eq. \ref{eq:wavenumber}, we expand Eq. \ref{eq:wave_action} to obtain:

\begin{equation}
\label{eq:wave_action2}
\dfrac{\partial N}{\partial t}
+ \left(C_g + U\right) \dfrac{\partial N}{\partial x}
+ N \dfrac{\partial U}{\partial x}
+ N \dfrac{\partial C_g}{\partial x}
= 0
\end{equation}
This equation is similar to Eq. \ref{eq:wavenumber}, except for the last term
which here represents the inhomogeneity of the group speed.
As \citet{peureux2021unsteady} explained, this term is responsible for unstable
growth of wave action in infinite long-wave trains.
Eqs. \ref{eq:wavenumber} and \ref{eq:wave_action2} are the governing equations
that we approximate and solve analytically in section \ref{section:analytical_solutions},
and solve numerically in their full form in section \ref{section:numerical_solutions}.

For the ambient forcing, we consider a train of monochromatic long wave
defined, to the first order in steepness, by the elevation $\eta_L = a_L cos{\psi}$,
where $\psi = k_L x - \omega_L t$ is the long-wave phase and $\omega_L$ its
angular frequency.
The long wave induces horizontal and vertical orbital velocities whose values
at the surface $\eta_L$ are:

@@

Eqs. \ref{eq:dispersion}-\ref{eq:w_L} is the full system of equations that we
use in this paper.
Note that here we derive the solutions from a different starting set of
equations than that of \citet{longuet1960changes}.
Rather than starting from the conservation of crests and wave action, as we do
here, they derived their solution from the linear superposition of the velocity
potential of two waves.
It appears that starting from the conservation of crests and wave action allows
for a simpler and more intuitive derivation.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/fig_hydrodynamic_modulation_diagram.pdf}
\caption{
  Diagram of the hydrodynamic modulation process.
}
\label{fig:hydrodynamic_modulation_diagram}
\end{figure}

@@

We now describe the analytical solutions for the short-wave wavenumber,
effective gravitational acceleration, amplitude, steepness, intrinsic frequency,
and intrinsic phase speed in presence of long waves.
The wavenumber is derived by linearizing the conservation of wave crests
(Eq. \ref{eq:wave_crests}) and neglecting the spatial gradients of $k$.
The effective gravity is derived by projecting the long-wave induced surface
accelerations onto the axis that is normal to the long-wave surface.
(Eq. \ref{eq:wave_action2}) and using the expression for the long-wave orbital
velocities.
The amplitude is derived by first evaluation the short-wave action from the
action conservation equation (Eq. \ref{eq:wave_action}) and then applying the
modulated effective gravity to obtain the amplitude.
From the modulated wavenumber, amplitude, and effective gravity, the steepness,
frequency, and phase speed readily follow from the usual wave relationships.
Going forward we will use the tilde over the variables to denote the modulated
quantities (e.g. $\widetilde{k}$ for wavenumber).

@@

\begin{equation}
\label{eq:wavenumber2}
\dfrac{\partial k}{\partial t}
= - k \dfrac{\partial U}{\partial x}
\end{equation}
Although \citet{peureux2021unsteady} pointed out that the solution by L-HS
requires assuming homogeneous group speed of the short-wave field, in fact,
it also requires assuming no horizontal propagation and advection of the short
waves.
Combining Eqs. \ref{eq:U_L} and \ref{eq:wavenumber2}, and integrating it in time
yields:

@@

Notice that the Taylor expansion of Eq. \ref{eq:k_short_exact} to the first order
recovers the original solution by \citet{longuet1960changes}:

\begin{equation}
\label{eq:k_short_lhs}
\frac{\widetilde{k}}{k} = 1 + \varepsilon_L \cos{\psi}
\end{equation}
Coming from the conservation of wave crests, as opposed to the velocity potential
superposition approach, the L-HS solution requires two approximations in their
solution:
First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
second, further truncating the Taylor expansion series beyond the first order term.
These approximations cause an underestimate of the short-wave modulation
magnitude (Fig. \ref{fig:analytical_solutions_ak0.1}a).
At the wave crest, Eq. \ref{eq:k_short_exact} yields the wavenumber modulation
that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of \citet{longuet1960changes}
for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
Although not immediately obvious from the expressions, the modulated wavenumber
in Eq. \ref{eq:k_short_lhs} is conservative across the long-wave phase, whereas
the higher-order solution in Eq. \ref{eq:k_short_exact} is not.
This is due to the fact that combining the orbital velocities evaluated at
$z=\eta$ and the linearized conservation of wave crests in Eq. \ref{eq:wavenumber}
results in excess short-wave wavenumber being created.
The solution of the full Eq. \ref{eq:wavenumber} of course must be conservative
across the long-wave phase, as we will see later from the numerical simulations.

@@

Long waves induce orbital accelerations, so the short waves that ride on their
surface move in an accelerated reference frame and experience effective
gravitational acceleration that is lower at the crests and higher in the troughs
\citep{longuet1986eulerian,longuet1987propagation}.
Further, centripetal force due to the curvature of the long wave induces small
but non-negligible effects on the short-wave effective gravity
\citep{phillips1981dispersion,zhang1990evolution}.
These effects are completely described by projecting the gravitational
acceleration from the coordinate that is perpendicular to the mean water surface
($z=0$) to that which is perpendicular to the long-wave surface ($z=\eta$),
and likewise for the long-wave orbital accelerations:

@@

\begin{equation}
\label{eq:local_slope}
\alpha = \tan^{-1}{\dfrac{\partial \eta}{\partial x}}
\end{equation}
where $\alpha$ is the local slope of the long-wave surface.
For a linear long wave, the effective gravity is:

\begin{equation}
\label{eq:gravity_modulation_linear}
\frac{\widetilde{g}}{g}
  = 
  \frac{
    1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}
    \left[ 1 + \left(\varepsilon_L \sin{\psi}\right)^2 \right]
  }
  {\sqrt{\left(\varepsilon_L \sin{\psi}\right)^2 + 1}}
\end{equation}

Curvilinear effects on effective gravity can be removed altogether by setting
$\alpha = 0$.
In that case, Eq. \ref{eq:gravity_modulation_general} simplifies to:

@@

The general form of the gravity modulation in Eq. \ref{eq:gravity_modulation_general}
is convenient because it allows us to easily attribute the effects of the different
processes on the gravity modulation.
It also makes it straightforward to evaluate the gravity modulation for a different
form of the long-wave elevation, by evaluating the orbital accelerations and the
local elevation slope at the appropriate level.
For a third-order Stokes wave, for example, the elevation is:

\begin{equation}
\label{eq:eta_stokes}
\eta_{St} = a_L \left[
  \cos{\psi} +
  \dfrac{1}{2} \varepsilon_L \cos{2\psi} +
  \varepsilon_L^2 \left( \dfrac{3}{8} \cos{3\psi} - \dfrac{1}{16} \cos{\psi} \right)
\right]
\end{equation}
Without considering the curvilinear effects in Eq. \ref{eq:gravity_modulation_general},,
the gravitational acceleration at the surface of a Stokes wave is, to the
fourth order:

\begin{equation}
\label{eq:gravity_modulation_stokes}
\frac{\widetilde{g}}{g} =
\left\{
  1 - \varepsilon_L e^{k \eta_{St}}
  \left[ \cos{\psi} -
    \varepsilon_L \sin{\psi} \left(
      \sin{\psi}
      + \varepsilon_L \sin{2\psi}
      - \dfrac{1}{16} \varepsilon_L^2 \sin{\psi}
      + \dfrac{9}{8} \varepsilon_L^2 \sin{3\psi}
    \right)
  \right]
\right\}
\end{equation}
The equivalent expression that includes the curvilinear effects can be derived
by evaluating orbital accelerations and local slope $\alpha$ at $z = \eta_{St}$,
and inserting them into Eq. \ref{eq:gravity_modulation_general}:

@@

Let us now examine the contributions of the different terms to the short-wave
effective gravity modulation.
Fig. \ref{fig:effective_gravities}a shows the effective gravity modulation of
short-wave as a function of the long-wave phase for $\varepsilon_L = 0.2$, for
a variety of the gravity modulation forms.
This value of $\varepsilon_L$ allows easier visualization of the modulation
differences, while still being relatively moderate.
To the first order in $\varepsilon_L$, the gravity modulation is out of phase
with the long-wave elevation, with the short waves experiencing a lower effective
gravity on the long-wave crests and a higher effective gravity in the troughs.
Evaluating the orbital accelerations at $z=\eta$ causes a gravity reduction
around the crests ($\approx6\%$) and the troughs ($\approx 3\%$), and an
increase on the front and rear faces ($\approx 4\%$) of the long wave, due to
the exponential change in the orbital accelerations away from the mean water level
(Fig. \ref{fig:effective_gravities}b).
The difference between the linear and Stokes long waves at this steepness is
negligible.
The long-wave elevation curvature has a second-order effect in $\varepsilon_L$.
Naturally, there is no curvature exactly at the crests and troughs, and its
effects are largest at the front and rear faces of the long wave, where the
covariance between the local slope and the horizontal accelerations are greatest.

@@

The modulation of short-wave amplitude can be derived in a similar way as we did
for the wavenumber, except that here we linearize the wave action balance and
integrate wave action in time:

>>>>

\begin{equation}
\label{eq:lhs1960wavenumber}
\widetilde{k} = k (1 + \varepsilon_L)
\end{equation}
where $\widetilde{k}$ and $k$ are the modulated and unmodulated short-wave wavenumbers,
respectively, and $\varepsilon_L = a_L k_L$ is the steepness of the long wave with
wavenumber $k_L$ and amplitude $a_L$.
This process is illustrated in Fig. \ref{fig:hydrodynamic_modulation_diagram}.
Hydrodynamic modulation is a key process for the development and
interpretation of remote sensing of ocean surface waves and currents
\citep{keller1975microwave,hara1994hydrodynamic}, and is distinguished
from aerodynamic modulation \citep{donelan1987effect,belcher1999wave,chen2000effects}.
\citet{phillips1981dispersion} and \citet{longuet1987propagation} revisited the
hydrodynamic modulation problem by considering nonlinear long waves and
variations in the effective gravity of short waves.
Their results yielded significantly stronger modulation than previously predicted.
\citet{henyey1988energy} derived an analytical solution for the modulation of
short waves using Hamiltonian mechanics and reported similar modulation magnitudes
to that of \citet{longuet1987propagation}.
\citet{zhang1990evolution} considered weakly nonlinear short waves
using a non-linear Schr\" odinger equation and found stronger wavenumber
modulation but weaker amplitude modulation than those predicted by
\citet{longuet1987propagation}.
The variety of analytical and numerical frameworks (Eulerian crest and action
conservation, Hamiltonian, Schr\" odinger) that relatively closely reproduce
steady solutions for short-wave modulation suggests that the solutions are robust.
Nevertheless, alternative analytical and numerical approaches to the modulation
problem remain of interest, especially if higher degrees of nonlinearity can be
considered using simpler approaches.

\begin{figure}
\centering
\includegraphics[width=\textwidth]{figures/fig_hydrodynamic_modulation_diagram.pdf}
\caption{
  A diagram of short waves riding on longer waves that propagate with their
  phase-speed $C_{pL}$ from left to right.
  The short waves propagate on the long-wave surface with their group velocity
  $C_g$.
  Long waves induce orbital velocities $U$ that cause near-surface convergence
  and divergence on their front and rear faces, respectively.
  They also induce downward and upward centripetal accelerations
  ($\partial W/\partial t$) in the crests and troughs, respectively, that
  modulate the effective gravity at the surface.
  As a consequence of surface convergence and divergence and the effective-gravity
  modulation, the short waves become shorter and higher preferentially
  on the long-wave crests.
  The diagram is not to scale.
}
\label{fig:hydrodynamic_modulation_diagram}
\end{figure}

All modulation solutions mentioned thus far are steady in the reference frame
of the long wave---short waves become shorter and higher (and thus, steeper)
preferentially on the long-wave crests.
Conversely, they elongate and flatten preferentially in the long-wave troughs.
Recently, \citet{peureux2021unsteady} asked whether the steady solutions
of short-wave modulation are appropriate by simulating the full wave conservation
equations.
They found that the short waves grow unsteadily due to the propagation of
longer waves, suggesting that the steady-state solutions may only be valid for
a few long-wave periods before the short waves begin to break due to excessive
steepness.
However, the unsteadiness of their solutions occurs in a specific scenario in
which a long-wave train suddenly appears to modulate a uniform field of short
waves.
The numerical simulations stabilize if the short-wave field is initialized
from existing steady solutions, like those by \citet{longuet1960changes}.
Nevertheless, their results do put into question whether the short-wave
modulation predicted by theory is generally steady over many long-wave periods
or not.
Looking at the conservation equations alone, it is not obvious that it is.

In this paper, we revisit this problem from linear wave theory but consider
the long-wave slope to be significant enough to require evaluating the
short-waves at the long-wave surface rather than at the mean water level.
This allows us to derive alternative, steady, non-linear solutions for the
modulation of short waves.
The analytical solutions yield modulations that are stronger than the original
solution by \citet{longuet1960changes}, but similar to the nonlinear numerical
solutions by \citet{longuet1987propagation} and \citet{zhang1990evolution} in
long-wave steepness up to $\approx 0.2$.
To validate the steadiness of the analytical solutions, we perform numerical
simulations of the full wave crest and action conservation laws, and compare
the results to the analytical solutions presented here, as well as the prior
numerical results.
The numerical simulations are used to determine the validity of the approximate
steady solution and to investigate the unsteadiness of the short-wave modulation
found by \citet{peureux2021unsteady}.
The main prior studies on hydrodynamic modulation, and their assumptions and
approaches to the solution, are summarized in Table
\ref{table:modulation_literature_summary}.
We first present the governing equations in \S\ref{section:governing_equations},
and then the analytical solutions for the modulation of short waves by long waves
in \S\ref{section:analytical_solutions}.
Then, we describe the numerical simulations and their results in
\S\ref{section:numerical_solutions}.
Finally, we discuss the results and their implications in
\S\ref{section:discussion} and conclude the paper in \S\ref{section:conclusions}.

\begin{table}
\begin{center}
\def~{\hphantom{0}}
\begin{tabular}{cccccc}
Reference   & Governing & Gravity & Nonlinear  & Nonlinear   & Solution \\
            & equations & varies  & long waves & short waves &          \\
\hline
\citet{longuet1960changes} & Velocity potential  & No  & No  & No  & Analytical \\
\hline
\citet{phillips1981dispersion} & Crest and action & Yes & Yes & No  & Numerical \\
                               & conservation     &     &     &     &           \\
\hline
\citet{longuet1987propagation} & Velocity potential & Yes & Yes & No  & Numerical \\
\hline
\citet{henyey1988energy} & Hamiltonian & Yes & Yes & No  & Analytical \\
\hline
\citet{zhang1990evolution} & Schr\" odinger & Yes & Yes & Yes & Numerical \\
\hline
\citet{peureux2021unsteady} & Crest and action & Yes & No  & No  & Numerical \\
                            & conservation     &     &     &     &           \\
\hline
This paper  & Crest and action & Yes & Yes & No  & Analytical \\
            & conservation     &     &     &     & and numerical \\
\hline
\end{tabular}
\caption{
  Summary of prior and present (this paper) approaches to calculating the
  modulation of short waves by long waves.
}
\label{table:modulation_literature_summary}
\end{center}
\end{table}

@@

The flow is inviscid, irrotational, and incompressible, and without sources or
sinks.
In these conditions, linear, deep-water, surface gravity waves obey the
dispersion relation:

\begin{equation}
\label{eq:dispersion}
\omega = \sqrt{gk} + k U
\end{equation}
where $\omega$ is the angular frequency in a fixed reference frame, $g$ is the
gravitational acceleration, $k$ is the wavenumber, and $U$ is the mean advective
current in the direction of the wave propagation.
A current in the direction of the waves increases their absolute frequency
without changing their wavenumber.
An important limitation here is that $U$ must be slowly varying on the scales of
the wave period \citep{bretherton1968wavetrains}.
From the perspective of short waves riding on long waves, the advective current
is the horizontal near-surface orbital velocity of the long wave.
Strictly, a velocity profile must be considered when determining the
advective velocity for a wave of any given wavenumber \citep{stewart1974hf},
however for brevity here we will assume a surface velocity to be the advective
one for short waves.
The evolution of the short-wave wavenumber is described by the conservation of
wave crests \citep{phillips1981dispersion}:

\begin{equation}
\label{eq:wave_crests}
\dfrac{\partial k}{\partial t}
+ \dfrac{\partial \omega}{\partial x}
= 0
\end{equation}
where $t$ and $x$ are time and space, respectively.
This relation states that the wavenumber must change locally when the frequency
varies in space.

(\ref{eq:dispersion}) and (\ref{eq:wave_crests}) only describe the evolution of
the short-wave wavenumber.
To describe the evolution of the wave amplitude, use the wave action balance
\citep{bretherton1968wavetrains} in absence of sources (e.g. due to wind) and
sinks (e.g. due to whitecapping):

\begin{equation}
\label{eq:wave_action}
\dfrac{\partial N}{\partial t}
+ \dfrac{\partial}{\partial x} \left[\left(C_g + U\right)N\right]
= 0
\end{equation}
where $N$ is the action density, and $C_g = 1/2\sqrt{g/k}$ is the group speed
in deep water.
Although the absence of wave growth and dissipation is not generally realistic
in the open ocean, it is a helpful approximation here to isolate the effects
of hydrodynamic modulation solely due to the motion of the long waves.

Short waves that ride on the surface of the longer waves move in an accelerated
reference frame (due to the centripetal acceleration at the long-wave surface)
and thus experience effective gravitational acceleration that varies in space
and time \citep{phillips1981dispersion,longuet1986eulerian,longuet1987propagation}.
Combine (\ref{eq:dispersion}) and (\ref{eq:wave_crests}) to get:

\begin{equation}
\label{eq:wavenumber}
\dfrac{\partial k}{\partial t}
+ \left(C_g + U\right) \dfrac{\partial k}{\partial x}
+ k \dfrac{\partial U}{\partial x}
+ \dfrac{1}{2} \sqrt{\dfrac{k}{g}} \dfrac{\partial g}{\partial x}
= 0
\end{equation}
From left to right, the terms in this equation represent the change in wavenumber
due to the propagation and advection by ambient current $U$, the convergence of
the ambient current, and the spatial inhomogeneity of the gravitational acceleration.
In a non-accelerated reference frame (i.e. in absence of longer waves), the last
term vanishes.
It is otherwise necessary to conserve the wave crests.

For the wave action, expand (\ref{eq:wave_action}) to get:

\begin{equation}
\label{eq:wave_action2}
\dfrac{\partial N}{\partial t}
+ \left(C_g + U\right) \dfrac{\partial N}{\partial x}
+ N \dfrac{\partial U}{\partial x}
+ N \dfrac{\partial C_g}{\partial x}
= 0
\end{equation}
This equation is similar to (\ref{eq:wavenumber}), except for the last term
which represents the inhomogeneity of the group speed.
As \citet{peureux2021unsteady} explained, the presence of this term causes
unstable growth of wave action in infinite long-wave trains.
(\ref{eq:wavenumber}) and (\ref{eq:wave_action2}) are the governing equations
that we approximate and solve analytically in \S\ref{section:analytical_solutions},
and numerically in their full form in \S\ref{section:numerical_solutions}.

For the ambient forcing, we consider a train of monochromatic long wave
defined, to the first order in steepness, by the elevation $\eta_L = a_L cos{\psi}$,
where $\psi = k_L x - \sigma_L t$ is the long-wave phase and $\sigma_L$ its
angular frequency.
The long wave induces horizontal and vertical orbital velocities whose values
at the surface $\eta_L$ are:

@@

(\ref{eq:dispersion})-(\ref{eq:w_L}) are the governing equations used in this paper.
Note that rather than deriving the modulation solution from the velocity
potential of two waves, as \citet{longuet1960changes} did, starting from the
crest-action conservation balances allows for a simpler derivation.

@@

We now describe the analytical solutions for the short-wave wavenumber,
effective gravitational acceleration, amplitude, steepness, intrinsic frequency,
and intrinsic phase speed in presence of long waves.
The wavenumber is derived by linearizing the conservation of wave crests
(\ref{eq:wave_crests}) and neglecting the spatial gradients of $k$.
The effective gravity is derived by projecting the long-wave induced surface
accelerations onto the local vertical axis (that is, the axis normal to the
long-wave surface), akin to \citet{zhang1990evolution}.
The amplitude is derived by first evaluating the short-wave action from the
action conservation (\ref{eq:wave_action}) and then applying the modulated
effective gravity to obtain the amplitude.
From the modulated wavenumber, amplitude, and effective gravity, the steepness,
frequency, and phase speed readily follow from the usual wave relationships.
Going forward we will use the tilde over the variables to denote the modulated
quantities (e.g. $\widetilde{k}$ for wavenumber).

@@

\begin{equation}
\label{eq:wavenumber2}
\dfrac{\partial k}{\partial t}
= - k \dfrac{\partial U}{\partial x}
\end{equation}
Although \citet{peureux2021unsteady} pointed out that the solution by
\citet{longuet1960changes} requires assuming homogeneous group speed of the
short-wave field, in fact, it also requires assuming no horizontal propagation
and advection of the short waves.
Combine (\ref{eq:U_L})-(\ref{eq:wavenumber2}), and integrate in time to get:

@@

Notice that the Taylor expansion of (\ref{eq:k_short_exact}) to the first order
recovers the original solution by \citet{longuet1960changes}:

\begin{equation}
\label{eq:k_short_lhs}
\frac{\widetilde{k}}{k} = 1 + \varepsilon_L \cos{\psi}
\end{equation}
Coming from the conservation of crests, as opposed to the velocity potential
superposition, the solution by \citet{longuet1960changes} requires two
approximations:
First, evaluating the long-wave velocity at $z = 0$ rather than $z = \eta_L$; and
second, further truncating the Taylor expansion series beyond the first-order
term.
These approximations underestimate the short-wave modulation magnitude
(Fig. \ref{fig:analytical_solutions_ak0.1}a).
At the wave crest, (\ref{eq:k_short_exact}) yields the wavenumber modulation
that is 16.9\%, 38.5\%, 66.4\%, and 104\% larger than that of
\citet{longuet1960changes} for $\varepsilon_L = 0.1, 0.2, 0.3, 0.4$, respectively.
Although not immediately obvious from the expressions, the modulated wavenumber
(\ref{eq:k_short_lhs}) is conserved across the long-wave phase, whereas
the higher-order solution (\ref{eq:k_short_exact}) is not.
This is because combining the orbital velocities evaluated at $z=\eta$ and the
linearized conservation of wave crests in (\ref{eq:wavenumber}) results in
excess short-wave wavenumber being created.
The solution of (\ref{eq:wavenumber}) must be conservative across the long-wave
phase, as we will see later from the numerical simulations.

@@

Long waves induce orbital (centripetal) accelerations on their surface, so the
short waves ride in an accelerated reference frame and experience effective
gravity that is lower at the crests and higher in the troughs
\citep{longuet1986eulerian,longuet1987propagation}.
Further, the projection of the long-wave induced horizontal acceleration
contributes to small but non-negligible effects on the short-wave effective
gravity in the curvilinear coordinate system
\citep{phillips1981dispersion,zhang1990evolution}.
These two effects are completely described by projecting the gravitational
acceleration vector from the coordinate that is perpendicular to the mean water
surface ($z=0$) to that which is perpendicular to the long-wave surface
($z=\eta_L$), and likewise for the long-wave orbital accelerations:

@@

\begin{equation}
\label{eq:local_slope}
\alpha = \tan^{-1}{\dfrac{\partial \eta}{\partial x}}
\end{equation}
where $\alpha$ is the local slope of the long-wave surface.
For short waves on a linear long wave, the gravity modulation is:

\begin{equation}
\label{eq:gravity_modulation_linear}
\frac{\widetilde{g}}{g}
  = 
  \frac{
    1 - \varepsilon_L \cos{\psi} e^{\varepsilon_L \cos{\psi}}
    \left[ 1 + \left(\varepsilon_L \sin{\psi}\right)^2 \right]
  }
  {\sqrt{\left(\varepsilon_L \sin{\psi}\right)^2 + 1}}
\end{equation}
Curvilinear effects on effective gravity can be removed altogether by setting
$\alpha = 0$.
In that case, (\ref{eq:gravity_modulation_general}) simplifies to:

@@

The general form of the gravity modulation (\ref{eq:gravity_modulation_general})
is convenient because it allows us to easily attribute the modulation to
different processes.
It also makes it straightforward to evaluate the gravity modulation for a
different long-wave form, by evaluating the orbital accelerations and the local
slope for that wave.
For a third-order Stokes wave, for example, the elevation is:

\begin{equation}
\label{eq:eta_stokes}
\eta_{St} = a_L \left[
  \cos{\psi} +
  \dfrac{1}{2} \varepsilon_L \cos{2\psi} +
  \varepsilon_L^2 \left( \dfrac{3}{8} \cos{3\psi} - \dfrac{1}{16} \cos{\psi} \right)
\right]
\end{equation}
Without considering the curvilinear effects in (\ref{eq:gravity_modulation_general}),
the gravitational acceleration at the surface of a Stokes wave is, to the third
order:

\begin{equation}
\label{eq:gravity_modulation_stokes}
\frac{\widetilde{g}}{g} =
\left\{
  1 - \varepsilon_L e^{k \eta_{St}}
  \left[ \cos{\psi} -
    \varepsilon_L \sin{\psi} \left(
      \sin{\psi}
      + \varepsilon_L \sin{2\psi}
      - \dfrac{1}{16} \varepsilon_L^2 \sin{\psi}
      + \dfrac{9}{8} \varepsilon_L^2 \sin{3\psi}
    \right)
  \right]
\right\}
\end{equation}
The equivalent expression that includes the curvilinear effects can be derived
by evaluating the velocities and the local slope $\alpha$ at $z = \eta_{St}$,
and inserting them into (\ref{eq:gravity_modulation_general}):

@@

Let us now examine the contributions of the different terms to the short-wave
effective gravity modulation.
Fig. \ref{fig:effective_gravities}a shows the effective gravity modulation of
short-wave as a function of the long-wave phase for $\varepsilon_L = 0.2$, for
a variety of the gravity modulation forms.
This value of $\varepsilon_L$ allows easier visualization of the modulation
differences, while still being relatively moderate.
To the first order in $\varepsilon_L$, the gravity modulation is out of phase
with the long-wave elevation, with the short waves experiencing a lower effective
gravity on the long-wave crests and a higher effective gravity in the troughs.
Evaluating the orbital accelerations at $z=\eta_L$ causes a gravity reduction
around the crests ($\approx6\%$) and the troughs ($\approx 3\%$), and an
increase on the front and rear faces ($\approx 4\%$) of the long wave, due to
the exponential change in the orbital accelerations away from the mean water level
(Fig. \ref{fig:effective_gravities}b).
The difference between the linear and Stokes long waves at this steepness is
negligible.
The long-wave elevation curvature has a second-order effect in $\varepsilon_L$.
Naturally, there is no curvature exactly at the crests and the troughs, and its
effects are largest at the front and rear faces of the long wave, where the
covariance between the local slope and the horizontal accelerations are largest.

@@

The modulation of short-wave amplitude can be derived in a similar way as we did
for the wavenumber, except that here we linearize the wave action balance
(\ref{eq:wave_action2}) and integrate it in time:
